\documentclass{ucbthesis}

% http://tug.ctan.org/info/biblatex-cheatsheet/biblatex-cheatsheet.pdf
\usepackage[giveninits=true, style=authoryear, hyperref=true, maxbibnames=8, maxcitenames=1, sorting=nyt, uniquename=init]{biblatex}
\usepackage{amsmath,amssymb,euscript,mathrsfs,amsthm}
\usepackage{hyperref}
\usepackage{microtype}
\usepackage{graphicx}
\usepackage{textcomp}
\usepackage{float}
%\usepackage{svg}
%\usepackage{subfigure}
\usepackage{caption}
\usepackage{subcaption}
\usepackage{amsfonts}
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{mathtools}
\usepackage{booktabs} % for professional tables


% Double spacing, if you want it.  Do not use for the final copy.
% \def\dsp{\def\baselinestretch{2.0}\large\normalsize}
% \dsp

% If the Grad. Division insists that the first paragraph of a section
% be indented (like the others), then include this line:
% \usepackage{indentfirst}

\newcommand{\dxdt} { \frac{dx}{dt} }
\newcommand{\dxdtn}[1] { \frac{d^{#1}x}{dt^{#1}}} % nth derivative
\newcommand{\ddt} { \frac{d}{dt} }
\newcommand{\argmin}{\operatornamewithlimits{argmin}}
\newcommand{\tightoverset}[2]{\mathop{#2}\limits^{\vbox to -.6ex{\kern-0.75ex\hbox{$#1$}\vss}}}

\DeclareCiteCommand{\citeyear}
    {\usebibmacro{prenote}}
    {\bibhyperref{\printfield{year}}\bibhyperref{\printfield{extrayear}}}
    {\multicitedelim}
    {\usebibmacro{cite:postnote}}

\DeclareCiteCommand{\citeyearpar}[\mkbibparens]
    {\usebibmacro{prenote}}
    {\bibhyperref{\printfield{year}}\bibhyperref{\printfield{extrayear}}}
    {\multicitedelim}
    {\usebibmacro{cite:postnote}}

\graphicspath{ {./figures/} }

\hyphenation{mar-gin-al-ia}
\hyphenation{bra-va-do}

\AtEveryBibitem{\clearfield{month}{} \clearlist{language}{}}
\DeclareNameAlias{sortname}{last-first}
\bibliography{references}

\begin{document}

% Declarations for Front Matter

\title{Analysis and applications of the Locally Competitive Algorithm}
\author{Dylan M Paiton}
\degreesemester{Spring}
\degreeyear{2019}
\degree{Doctor of Philosophy}
\chair{Professor Bruno Olshausen}
\othermembers{Professor Jack Gallant\\
  Professor Stanley Klein\\
  Professor Trevor Darrell}
\numberofmembers{4}
% Previous degrees are no longer to be listed on the title page.
% \prevdegrees{B.A. (University of Northern South Dakota at Hoople) 1978 \\
%   M.S. (Ed's School of Quantum Mechanics and Muffler Repair) 1989}
\field{Vision Science}
% Designated Emphasis -- this is optional, and rare
% \emphasis{Colloidal Telemetry}
% This is optional, and rare
% \jointinstitution{University of Western Maryland}
% This is optional
\campus{Berkeley}

% For a masters thesis, replace the above \documentclass line with
% \documentclass[masters]{ucbthesis}
% This affects the title and approval pages, which by default calls this
% document a "dissertation", not a "thesis".

\maketitle
% Delete (or comment out) the \approvalpage line for the final version.
%\approvalpage
\copyrightpage

\include{abstract}

\begin{frontmatter}

\begin{dedication}
\null\vfil
\begin{center}
Dedication.\\\vspace{12pt}
I would like to dedicate this thesis to Janet Paiton, Donnie Paiton, Loni Hammer, Justin Hammer, and Christine Hopkins. Your love and support have kept my spirits high and my life on track, and I can't imagine having done any of this without you. I hope to continue to make you proud and reward the valuable life lessons you gave me. Thank you all so much!
\end{center}
\vfil\null
\end{dedication}

% You can delete the \clearpage lines if you don't want these to start on
% separate pages.

\tableofcontents
\clearpage
\listoffigures
\listoftables

\begin{acknowledgements}
This thesis benefited greatly from feedback and contribution from my colleagues and friends. Christine Hopkins, Wesley Chavez, Will Shainin, Sheng Lundquist, Gar Kenyon, Matt Brown, and Chris Warner all read portions of this text and provided valuable feedback. Thanks Vasha Dutell for helping me hunt down sources. Thanks Parker Conroy for pushing me to make continual progress. The hard-thresholded LCA analysis work was done with Will Shainin and Gar Kenyon. The weakly supervised learning work was done under the advisement of Jack Culpepper at Yahoo, Inc. Simon Osindero, Rob Hess, Clayton Mellina, and Tobi Baumgartner also provided valuable feedback on that project. The subspace LCA work was done in collaboration with Ryan Chan. The Iso-Response contour work was done in collaboration with Charles Frye, Joel Bowen, Jasmine Collins, Sheng Lundquist, and Melanie Mitchell. The orientation selectivity experiments and rate-distortion experiments were done in collaboration with Sophia Sanborn and Spencer Kent. Also thank you Ryan Zarcone for being an invaluable soundboard and problem solver. Thank you to Vision Science faculty Austin Roorda, Stan Klein, Michael Silver, and Martin Banks for your support and friendship. Thank you to my thesis committee Bruno Olshausen, Stan Klein, Jack Gallant, and Trevor Darrell for your feedback.

I would like to thank the following for repeatedly taking their valuable time to work through challenges (both scientific and personal) with me: Charles Frye, Ryan Zarcone, Jesse Livezey, and Shariq Mobin. I am extremely grateful to have had an opportunity to work with such a collaborative group at the Redwood Center. Thank you to everyone in this list for listening, helping me sculpt my ideas, and providing feedback: Alex Anderson, Alex Terekhov, Brian Cheung, Bruno Olshausen, Charles Frye, Charles Garfinkle, Chris Hillar, Chris Warner, Connor Bybee, Eric Dodds, Eric Weiss, Friedrich Sommer, Gautam Agarwal, Guy Isley, James Arnemann, Jascha Sohl-Dickstein, Jasmine Collins, Jesse Engel, Jesse Livezey, Karl Zipser, Kris Bouchard, Louis Kang, Mayur Mudigonda, Michael Fang, Mike DeWeese, Neha Wadia, Paxon Frady, Pentti Kanerva, Pratik Sachdeva, Ryan Chan, Ryan Zarcone, Saeed Saremi, Shariq Mobin, Sophia Sanborn, Spencer Kent, Stephen Shepard, Tyler Lee, Urs Koster, Vasha Dutell, and Yubei Chen.

My foundation in theoretical neuroscience was established while working with the Garmy under the advisement of Gar Kenyon. Thank you so much Gar for continuing to provide invaluable support, advisement, and friendship. Thank you also to Pete Schultz, Steven Brumby, Amy Larson, John George, Michael Ham, Gerd Kunde, Brennan Nowers, and John Galbraith for providing project feedback and collaboration.

My sanity and happiness are \textit{strongly} dependent on support form my Very Best Friends and my Vision Science cohort. You all know who you are. You keep me going - thank you.

Finally, all of the work in this thesis was overseen by my amazing PhD advisor, Bruno Olshausen. He was a continual source of direction, encouragement, and inspiration for my work. It was his writing that pulled me into this field in the first place, and his passion for understanding neural computation has continued to stoke my own. Thank you, Bruno.
\end{acknowledgements}

\end{frontmatter}

\pagestyle{headings}

% (Optional) \part{First Part}

\include{chap1}\label{ch:intro}

\include{chap2}\label{ch:lca}

\include{chap3}\label{ch:hierarchical_sc}

\include{chap4}\label{ch:iso}

\include{chap5}\label{ch:conclusion}

%\appendix
%\chapter{appendix}

\clearpage
\printbibliography

\end{document}


% title - locally competitive algorithms
% deemphasize LCA as a single algorithm, instead it's a family
% the LCA -> LCA?

